Visual Diagnostics of an Explainer Model -- Tools for the Assessment of LIME Explanations

### 2021-01-06 paper.Rnw
small edit

% Citation information
\jnlcitation{
\cname{\author{Goode K.}, \author{H. Hofmann}},
\cyear{2020},
\ctitle{Visual Diagnostics of an Explainer Model -- Tools for the Assessment of LIME Explanations},
\cjournal{Stat Anal Data Min: The ASA Data Sci Journal},
\cvol{volume, number and page}.}

@@

@@

@
\caption{Hypothetical scenarios depicting the effect of different weights on LIME. (Left) An example of an explainer model with appropriate weights to provide a faithful approximation of the complex model. (Right) An example of an explainer model for the same prediction but with different weights that results in an explanation unfaithful to the complex model.}
\label{fig:figure-01}
\end{figure*}

In the field of statistics, there are two main uses for models: inference and prediction. Machine learning models have proven to perform well in problems with the latter objective, but the accuracy of many machine learning models comes at the cost of interpretability due to their algorithmic complexity (hence the phrase "black-box models"). Model interpretability allows for the understanding and assessment of how a model produces predictions. The lack of the ability to understand and assess a model makes it difficult to trust the model, especially in areas with high stakes decisions such as the medical and forensics sciences. The increased use of machine learning models in applications and the introduction of the General Data Protection Regulation (GDPR) in 2018 \citep{goodman:2016} has resulted in a dramatic increase in explainable machine learning research, which focuses on developing ways to explain output from machine learning algorithms.

Throughout this paper, we distinguish between interpretability and explanability of models. We define {\it interpretability} as the ability to directly use model parameters to understand the relationships in the data captured by the model: e.g., a linear model coefficient associated with a predictor variable indicates the amount the response variable changes based on a change in the predictor\sout{ variable}. In contrast, we define {\it explanability} as the ability to use the model in an indirect manner to understand the relationships in the data captured by the model: e.g., partial dependence plots depict the marginal relationship between model predictions and predictor variables \citep{friedman:2001}.

Numerous methods have been proposed to provide explanations for black-box model predictions \citep{gilpin:2018, guidotti:2018, ming:2017, molnar:2019}. Some are specific to one type of model (e.g. \citep{simonyan:2013} and \citep{urbanek:2008}), and others are model-agnostic (e.g. \citep{fisher:2018} and \citep{strumbelj:2014}). In this paper, we focus on the  model-agnostic method of LIME \citep{ribeiro:2016}.

LIME (local interpretable model-agnostic explanations) uses a surrogate model to relate predictor variables to black-box model predictions  \citep{ribeiro:2016}. While some explainer models focus on understanding a model at the global level, LIME claims to provide explanations for individual predictions (local). Additionally, LIME is designed to work with any model (model-agnostic) and to produce easily understandable results (explanations) \citep{ribeiro:2016}. Conceptually, LIME fits a simple (interpretable) model, the explainer model, to approximate  the complex model in a local region around a prediction of interest. The simple model is interpreted to identify the variables that most influence the complex model prediction.

@@

To demonstrate the effect of a tuning parameter on LIME, we consider two different explainer models applied to the same prediction. \autoref{fig:figure-01} depicts two plots of the predictions from a hypothetical black-box model versus the feature used to train the model. The location of a prediction of interest is indicated by the diamond shaped points. Globally, there is a clear non-linear relationship between the predictions and the feature, but the relationship could be approximated by a linear model in a local region around the prediction of interest. In both scenarios, a linear regression weighted by the proximity from an observation to the prediction of interest is used as the explainer model, but the proximities are computed differently as depicted by the varying sizes of the observations in the two plots.

On the left, the distance between an observation and the prediction of interest is computed using the Gower distance metric \citep{gower:1971} and then raised to the power of \Sexpr{good_gower_power} to emphasize a very local region. The distances are subtracted from 1 to obtain a proximity. Here, the explainer model is faithful to the complex model since it captures the linear relationship between the black-box predictions in an immediate neighborhood of the prediction of interest. The slope explains that the black-box predictions increase as the feature increases in the local region around the prediction of interest.

For the scenario on the right, the computation of the proximities is the same except the distances are raised to a power of \Sexpr{bad_gower_power} (the default exponent in the \emph{lime} R package \citep{pedersen:2020}). This causes the observations that are further away from the prediction of interest to be given larger weights than the scenario on the left. Here, the explainer model does not capture the linear trend near the prediction of interest, but instead, \sout{the explainer captures} a more global trend. Without an assessment of this explainer model, the slope incorrectly suggests that the predictions decrease as the feature increases near the prediction of interest.

\autoref{fig:figure-01} demonstrates a simple example in which the explanation quality varies drastically when different weights are used. Other concerns \sout{with} LIME have been raised in the literature: \citet{laugel:2018} and \citet{molnar:2019} discuss the difficulty specifying a local region with LIME due to both an unclear definition of \sout{what is} a "local region" and how to apply LIME to achieve an appropriate local region. \citet{laugel:2018} also present examples of LIME explanations that are clearly global and not local. A different concern about LIME is raised by \citet{alvarezmelis:2018} pertaining to the robustness of explanations from LIME and other explainer models: they find that even small changes in predictor variables can lead to very different LIME explanations. Even the original authors of LIME, \citet{ribeiro:2016}, acknowledge that if a linear model is used as the explainer, LIME relies on a linear approximation of the explainer model to the complex model and state "if the underlying model is highly non-linear even in the locality of the prediction, there may not be a faithful explanation".

Without an assessment of the LIME explainer model, the user is putting trust in another black-box \kge{algorithm} \sout{procedure} to explain the complex black-box model. In this paper, we stress the importance of assessing the LIME explainer model. To do this, we lay out the set of claims about LIME made by \citet{ribeiro:2016}, and we provide a toolkit of visual diagnostics  for the assessment of these claims: (1) \emph{explanation scatterplots}, (2) \emph{feature heatmaps}, (3) \emph{assessment metric plots}. Using these visual diagnostics, we demonstrate problematic examples of LIME explanations where the claims about LIME are not met and the explanations are highly dependent on tuning parameter values without a clear way to identify a set of tuning parameter values that produce the "best" explanations.

@@

\citet{ribeiro:2016} provide an implementation of LIME in a Python package \cite{ribeiro:2020}. An adaption of the Python package in R has been implemented and made available by \citet{pedersen:2020}. In both implementations, a ridge regression model is used as the explainer model, and the user \sout{is asked to} select\kge{s} various tuning parameter options: the number of features to return in the explanation, the simulation method, the feature selection method, and how the weights are computed. \sout{The discussion of parameter choices and implementation details of LIME in this paper are based on the R package.} See Appendix~\ref{lime-details} for an overview of the options available for the tuning parameters \kge{based on the R package}.

@@

# Save the training and testing data
if(!file.exists("data/sine-data-train.rds")) saveRDS(sine_data_train, "data/sine-data-train.rds")
if(!file.exists("data/sine-data-test.rds")) saveRDS(sine_data_test, "data/sine-data-test.rds")

@@

@
\caption{Plots of $x_2$ versus $x_1$ from the \data \ training (left) and testing (right) sets introduced in Section \ref{diagnostics} with the true classification boundary shown as a solid black line. In the testing set, the open black circles identify cases misclassified by the random forest, and the diamond indicates a prediction of interest.}
\label{fig:figure-02}
\end{figure*}

@@

@
\caption{(Left) Visualization from the \emph{lime} R package of a LIME explanation for the \data \ prediction of interest identified in \autoref{fig:figure-02}. (Right) \emph{Explanation Scatterplot}.  An explanation scatterplot associated with the explanation on the left. The four quantile bins over-simplify the relationship between the random forest predictions and the $x_1$ and $x_2$ values near the prediction of interest.}
\label{fig:figure-03}
\end{figure*}

@@

To demonstrate the visual diagnostics, we generate an example dataset that will be referred to as the \data. The \data \ contains \Sexpr{dim(sine_data)[1]} observations with  features of $x_1$ and $x_2$ independently sampled from Unif(\Sexpr{min}, \Sexpr{max}) distributions and $x_3$ sampled from a $\mbox{N}(0,1)$ distribution. A binary response variable $y$ is created using a sine curve such that
\begin{eqnarray}\label{eq:data}
  y=\begin{cases}
  \mbox{blue} & \mbox{ if } x'_2 > \sin\left(x'_1\right) \\
  \mbox{red} & \mbox{ if } x'_2 \le \sin\left(x'_1\right) \ .
  \end{cases}
\end{eqnarray}
where $x'_1=x_1\cos(\theta)-x_2\sin(\theta)$,  $x'_2=x_1\sin(\theta)+x_2\cos(\theta)$, and  $\theta=-0.9$. Note that due to the creation of $y$ in this manner, $y$ is dependent on $x_1$ and $x_2$ and independent of $x_3$. The dataset is randomly divided into training and testing sets of \Sexpr{dim(sine_data_train)[1]} and \Sexpr{dim(sine_data_test)[1]} observations, respectively. A random forest is fit to the training set using the R package \emph{randomForest} (version \Sexpr{packageVersion("randomForest")}) \citep{liaw:2002} with the default settings and is applied to the test set to obtain predictions.

@@

\sout{Since 4-quantile-bins} \kge{When quantile bins} are used to simulate data, LIME converts continuous predictor variables to indicator variables identifying whether the variable value falls in the same quantile bin as the prediction of interest or not. The indicator variables are used as the features in the ridge regression explainer model. The features included in the visualization are the ones selected by feature selection, the lengths of the bars represent the coefficients from the ridge regression associated with the indicator variables, while the color of a bar denotes the sign of the coefficient. This explainer model suggests that a random forest prediction of `blue' for this observation is mostly supported by the prediction of interest having a value of $x_2$ that is greater than \Sexpr{str_split(sine_poi_lime_default$feature_desc[[1]], pattern = " ")[[1]][1]} but somewhat contradicted by the prediction of interest having a value of $x_1$ that is greater than \Sexpr{str_split(sine_poi_lime_default$feature_desc[[2]], pattern = " ")[[1]][1]} and less than or equal to \Sexpr{str_split(sine_poi_lime_default$feature_desc[[2]], pattern = " ")[[1]][5]}. %\kgc{I decided to leave the values here. I actually prefer it this way myself, but I cut out a lot of text from this subsection in other ways.}\hh{XXX }

@@

For a further assessment of the explainer model, we turn to an \emph{explanation scatterplot}: a visual diagnostic for assessing the LIME claims of locality and fidelity for an individual explanation by juxtaposing the complex and explainer models in one plot. The format of an explanation scatterplot depends on the LIME simulation method. We introduce the explanation scatterplot here under the \emph{lime} R package default method of 4-quantile-bins. \kge{See the supporting information for} \sout{Appendix \ref{exp-scatter_plus} addresses the} explanation scatterplot formats \kge{under} \sout{in} other LIME simulation scenarios.

The explanation scatterplot applies the concept of plotting the model in the data space discussed in \citet{wickham:2015}. The \kge{visualization} \sout{plot} is built by plotting the LIME simulated data for the top two features identified by the explanation in a scatterplot and coloring these points by the predictions from the complex model. The point size represents the weight assigned by LIME. In order to show the LIME results for the observation of interest, lines are drawn on top of the points representing the boundaries of the indicator variables used to fit the explainer model. The line color denotes whether LIME indicates that a feature supports or contradicts a class prediction.

@@

First, consider the claim of localness. The weights decay relatively slowly outside of the intersection of the \kge{two quantile bins} \sout{2-quantile-bin} suggesting that the LIME explanation is highly influenced by points outside of the bins containing the prediction of interest. However, it is difficult to say if the claim of localness has been violated, because a definition of a local region is not specified. Depending on what region a viewer considers to be local, an argument could be made in favor of or against a violation of localness. \sout{The explanation scatterplot raises awareness of the unclear definition of a local region with LIME. Nevertheless, it is possible to say that the weights assigned to the simulated data do not capture the local region identified by the intersection of the quantile bins.}

Now, consider the claim of faithfulness: It can be said that the majority of the points in the $x_2$ quantile \kge{bin} support a prediction of blue, which is captured by the bar supporting a prediction of blue in the LIME explanation, and a similar statement can be made about the $x_1$ quantile-bin. These statements validate the explanation produced by LIME. However, the explanation scatterplot plot shows that the random forest performs well at capturing the sine curve classification boundary by creating various sized rectangles consisting of predictions with similar probabilities, which the LIME explanation does not pick up on since the bins are created without information about the random forest predictions. Thus, LIME does provide an explanation for the prediction, but it is a very poor explanation in terms of faithfulness to the random forest prediction regions.

@@

This example explanation \kge{scatterplot} only includes two features. In situations where more than two features are included in a LIME explanation, the explanation scatterplots can be extended to a generalized pairs plot \citep{emerson:2013} that includes all pairwise combinations of features. Generalized pairs plots (and scatterplot matrices in general) have diminishing value when the number of features increase \citep{jensen:2011} \citep{sweller:2011}. Machine learning models are commonly fit using a large number of features, and therefore, a generalized pairs plot of explanation scatterplots for all features would be ineffective. However, when applying LIME, the user selects the number of features to return in the explanation. In the \emph{lime} R package, \citet{pedersen:2020} encourage users to select less than 10 features. As long as a small number of features are returned in the LIME explanation, it is feasible to use a generalized pairs plot of explanation scatterplots. An example is shown in Section~\ref{bullet-assess-ex}.

@@

@
\caption{Example bullet signatures. The signatures are from the same land and therefore have very similar patterns.}
\label{fig:figure-07}
\end{figure}

@@

\begin{figure*}[!thp]
<<figure-11, out.width = '6.25in', fig.width = 13, fig.height = 9, message = FALSE>>=

@@

>>>>

% Citation information
\jnlcitation{
\cname{\author{Goode K.}, \author{H. Hofmann}, }
\cyear{2020},
\ctitle{Visual Diagnostics of an  Explainer Model -- Tools for the Assessment of LIME Explanations},
\cjournal{Stat Anal Data Min: The ASA Data Sci Journal},
\cvol{volume, number and page}.}

@@

<<concept-data, eval = FALSE>>=

# # Simulate example data
# set.seed(20190624)
# lime_data <-
#   data.frame(
#     feature1 = sort(runif(250, 0, 1)),
#     feature2 = sample(x = 1:250, size = 250, replace = FALSE)) %>%
#   mutate(
#     feature1_stnd = (feature1 - mean(feature1)) / sd(feature1),
#     feature2_stnd = (feature2 - mean(feature2)) / sd(feature2),
#     prediction = if_else(feature1 >= 0 & feature1 < 0.1,
#                          (0.3 * feature1) + rnorm(n(), 0, 0.01),
#                  if_else(feature1 >= 0.1 & feature1 < 0.3,
#                          rbeta(n(), 1, 0.5),
#                  if_else(feature1 >= 0.3 & feature1 < 0.5,
#                          sin(pi* feature1) + rnorm(n(), 0, 0.5),
#                  if_else(feature1 >= 0.5 & feature1 < 0.8,
#                          -(sin(pi* feature1) + rnorm(n(), 0, 0.1)) + 1,
#                  if_else(feature1 >= 0.8 & feature1 < 0.9,
#                          0.5 + runif(n(), -0.5, 0.5),
#                          0.5 + rnorm(n(), 0, 0.3))))))) %>%
#   bind_rows(
#     data.frame(feature1 = rep(1, 30) + rnorm(5, 0, 0.05),
#            feature2 = rep(245, 30) + rnorm(5, 0, 1),
#            prediction = rep(0, 30) + rnorm(5, 0, 0.05)) %>%
#   mutate(feature1_stnd = (feature1 - mean(feature1)) / sd(feature1),
#          feature2_stnd = (feature2 - mean(feature2)) / sd(feature2)))
# 
# # Specify a prediction of interest
# prediction_of_interest <-
#   data.frame(feature1 = 0.07, feature2 = 200) %>%
#   mutate(feature1_stnd = (feature1 - mean(lime_data$feature1)) /
#            sd(lime_data$feature1),
#          feature2_stnd = (feature2 - mean(lime_data$feature2)) /
#            sd(lime_data$feature2),
#          prediction = 0.05,
#          color = factor("Prediction \nof Interest"))
# 
# # Specify the gower exponents
# good_gower_power <- 50
# bad_gower_power <- 1
# 
# # Compute the good distances between the prediction of interest
# # and all other observations
# lime_data$distance_good <-
#   (1 - gower_dist(x = prediction_of_interest %>%
#                     select(feature1_stnd, feature2_stnd),
#                   y = lime_data %>%
#                     select(feature1_stnd,
#                            feature2_stnd)))^good_gower_power
# 
# # Compute the bad distances between the prediction of interest
# # and all other observations
# lime_data$distance_bad <-
#   (1 - gower_dist(x = prediction_of_interest %>%
#                     select(feature1_stnd, feature2_stnd),
#                   y = lime_data %>%
#                     select(feature1_stnd,
#                            feature2_stnd)))^bad_gower_power
# 
# # Prepare the data for plotting
# lime_data_gathered <- lime_data %>%
#   gather(feature, feature_stnd_value,
#          feature1_stnd:feature2_stnd) %>%
#   gather(distance, distance_value, distance_good:distance_bad) %>%
#   mutate(distance = fct_recode(distance,
#                                "good" = "distance_good",
#                                "bad" = "distance_bad"),
#          feature = fct_recode(feature,
#                               "Feature 1" = "feature1_stnd",
#                               "Feature 2" = "feature2_stnd"))
# 
# # Prepare the prediction of interest data for plotting
# prediction_of_interest_gathered <- prediction_of_interest %>%
#   select(-feature1, -feature2) %>%
#   gather(feature, feature_value, feature1_stnd, feature2_stnd) %>%
#   mutate(feature = fct_recode(feature,
#                               "Feature 1" = "feature1_stnd",
#                               "Feature 2" = "feature2_stnd"))

<<concept-explainers, eval = FALSE>>=

# # Fit the good interpretable explainer model
# explainer_good <-
#   glmnet(x = lime_data %>% select(feature1_stnd, feature2_stnd)
#          %>% as.matrix(),
#          y = lime_data$prediction,
#          alpha = 0,
#          lambda = 1,
#          weights = lime_data$distance_good)
# 
# # Fit the bad interpretable explainer model
# explainer_bad <-
#   glmnet(x = lime_data %>% select(feature1_stnd, feature2_stnd)
#          %>% as.matrix(),
#          y = lime_data$prediction,
#          alpha = 0,
#          lambda = 1,
#          weights = lime_data$distance_bad)
# 
# # Join the coefficients from the explainer model into a dataframe
# coefs_data <- data.frame(case = c("good", "bad"),
#                          b0 = c(coef(explainer_good)[1],
#                                 coef(explainer_bad)[1]),
#                          b1 = c(coef(explainer_good)[2],
#                                 coef(explainer_bad)[2]),
#                          b2 = c(coef(explainer_good)[3],
#                                 coef(explainer_bad)[3])) %>%
#   mutate(feature1 = b0 + b2*prediction_of_interest$feature2_stnd,
#          feature2 = b0 + b1*prediction_of_interest$feature1_stnd) %>%
#   gather(key = feature, value = int, feature1:feature2) %>%
#   mutate(slope = c(coef(explainer_good)[2],
#                    coef(explainer_good)[3],
#                    coef(explainer_bad)[2],
#                    coef(explainer_bad)[3]),
#          feature = fct_recode(feature,
#                               "Feature 1" = "feature1",
#                               "Feature 2" = "feature2"))

@@

# # Plot of good explainer model
# ggplot() +
#   facet_grid(. ~ feature, switch = "x") +
#   geom_point(
#     data = lime_data_gathered %>% mutate(
#       feature = fct_recode(
#         feature,
#         "Feature 1 (standardized)" = "Feature 1",
#         "Feature 2 (standardized)" = "Feature 2"
#       )
#     ),
#     mapping = aes(
#       x = feature_stnd_value,
#       y = prediction,
#       size = distance_value,
#       color = distance
#     ),
#     alpha = 0.6
#   ) +
#   geom_abline(
#     data = coefs_data %>%
#       filter(case == "good") %>%
#       mutate(
#         feature = fct_recode(
#           feature,
#           "Feature 1 (standardized)" = "Feature 1",
#           "Feature 2 (standardized)" = "Feature 2"
#           )
#         ),
#     mapping = aes(intercept = int, slope = slope),
#     size = 1
#   ) +
#   scale_color_manual(values = c(NA, "grey30"), guide = "none") +
#   geom_point(
#     data = prediction_of_interest_gathered %>% mutate(
#       feature = fct_recode(
#         feature,
#         "Feature 1 (standardized)" = "Feature 1",
#         "Feature 2 (standardized)" = "Feature 2"
#       )
#     ),
#     mapping = aes(x = feature_value,
#                   y = prediction,
#                   fill = color),
#     color = "black",
#     size = 5,
#     shape = 23,
#     alpha = 0.75
#   ) +
#   scale_fill_manual(values = "#FAAA72") +
#   geom_text(
#     data = data.frame(
#       feature_stnd_value = 1.2,
#       prediction = 1.6,
#       feature = c("Feature 1 (standardized)", "Feature 2 (standardized)"),
#       slope = c(
#         paste(
#           "Slope:",
#           coefs_data %>%
#             filter(case == "good",
#                    feature == "Feature 1") %>%
#             pull(slope) %>%
#             round(3)
#         ),
#         paste(
#           "Slope:",
#           coefs_data %>%
#             filter(case == "good",
#                    feature == "Feature 2") %>%
#             pull(slope) %>%
#             round(3)
#         )
#       )
#     ),
#     mapping = aes(x = feature_stnd_value,
#                   y = prediction,
#                   label = slope),
#     family = ff,
#     size = f1_fs / 2.85
#   ) +
#   labs(
#     x = "",
#     y = "Black-Box Prediction",
#     title = "Conceptual Depiction of a Faithful Local Explainer Model",
#     subtitle = paste("Gower Distance Metric Exponent:", good_gower_power),
#     fill = "",
#     alpha = "Weight",
#     size = "Weight"
#   ) +
#   theme_linedraw(base_family = ff, base_size = f1_fs) +
#   theme(
#     panel.grid.major = element_blank(),
#     panel.grid.minor = element_blank(),
#     axis.text.x = element_blank(),
#     axis.ticks.x = element_blank(),
#     axis.text.y = element_blank(),
#     axis.ticks.y = element_blank(),
#     strip.placement = "outside",
#     strip.background = element_rect(color = "white", fill = "white"),
#     strip.text.x = element_text(size = f1_fs, color = "black"),
#     plot.title = element_text(size = f1_fs)
#   ) +
#   guides(
#     fill = guide_legend(order = 1),
#     size = guide_legend(order = 2),
#     alpha = guide_legend(order = 2)
#   )

@
\caption{Hypothetical scenarios depicting the effect of different weights on LIME. (Left) An example of an explainer model with appropriate weights to provide a faithful approximation of the complex model. (Right) An example of an explainer model for the same prediction but with different weights that results in an explanation unfaithful to the complex model.  }
\label{fig:figure-01}
\end{figure*}

In the field of statistics, there are two main uses for models: inference and prediction. Machine learning models  have proven to perform well in problems with the latter objective, but the accuracy of many machine learning models comes at the cost of interpretability due to their algorithmic complexity (hence the phrase "black-box models"). Model interpretability allows for the understanding and assessment of how a model produces predictions. The lack of the ability to understand and assess a model makes trusting it difficult, especially in areas with high stakes decisions such as the medical and forensics sciences. The increased use of machine learning models in applications and the introduction of the General Data Protection Regulation (GDPR) in 2018 \citep{goodman:2016} has resulted in a dramatic increase in explainable machine learning research, which focuses on developing ways to explain output from machine learning algorithms.

Throughout this paper, we distinguish between interpretability and explanability of models. We define {\it interpretability} as the ability to directly use model parameters to understand the relationships in the data captured by the model: e.g., a linear model coefficient associated with a predictor variable indicates the amount the response variable changes based on a change in the predictor variable. In contrast, we define {\it explanability} as the ability to use the model in an indirect manner to understand the relationships in the data captured by the model: e.g., partial dependence plots depict the marginal relationship between model predictions and predictor variables \citep{friedman:2001}.

Numerous methods have been proposed to provide explanations for black-box model predictions \citep{gilpin:2018, guidotti:2018, ming:2017, molnar:2019}. Some are specific to one type of model (e.g. \citep{simonyan:2013, urbanek:2008}), others are model-agnostic (e.g. \citep{fisher:2018, strumbelj:2014}). In this paper, we focus on the  model-agnostic method of LIME \citep{ribeiro:2016}.

LIME (local interpretable model-agnostic explanations)  uses a surrogate model to relate predictor variables to black-box model predictions  \citep{ribeiro:2016}.  While some explainer models  focus on understanding a model at the global level, LIME claims to provide explanations for individual predictions (local). Beyond that, LIME is designed to work with any model (model-agnostic) and to produce easily understandable results (explanations) \citep{ribeiro:2016}. Conceptually, LIME fits a simple (interpretable) model, the explainer model, to approximate  the complex model in a local region around a prediction of interest. The simple model is interpreted to identify the  variables that most influence the complex model prediction.

@@

To demonstrate the effect of a tuning parameter on LIME, we consider two different explainer models applied to the same prediction. \autoref{fig:figure-01} depicts two plots of the predictions from a hypothetical black-box model versus the feature used to train the model. The location of a prediction of interest is indicated by the diamonds. Globally, there is a clear non-linear relationship between the predictions and the feature, but the relationship could be approximated by a linear model in a local region around the prediction of interest. In both scenarios, a linear regression weighted by the proximity from an observation to the prediction of interest is used as the explainer model, but the proximities are computed differently as depicted by the varying sizes of the observations in the two plots.

On the left, the distance between an observation and the prediction of interest is calculated using the Gower distance metric \citep{gower:1971}  raised to the power of \Sexpr{good_gower_power} to emphasize a very local region. The distances are subtracted from 1 to obtain a proximity. Here, the explainer model is faithful to the complex model since it captures the linear relationship between the black-box predictions in an immediate neighborhood of the prediction of interest: with an increase in the feature the black-box predictions increase in a local region around the prediction of interest.

For the scenario on the right, the computation of the proximities is the same except that distances are raised to a power of \Sexpr{bad_gower_power} (the default exponent in the \emph{lime} R package \citep{pedersen:2020}). This difference gives larger weights to observations that are further away from the prediction of interest than the scenario on the left. Instead of capturing a local trend, the slope in the explainer model describes a more global trend. Here, the explainer model incorrectly suggests, that a decrease in prediction values as the feature increases near the prediction of interest. %\sout{This cannot be said of the explainer model in \autoref{fig:figure-02}. In addition, the magnitude of the slope associated with Feature 2  (\Sexpr{coefs_data %>% filter(case == "bad", feature == "Feature 2") %>% pull(b1) %>% round(3) %>% abs()}) is larger than that of Feature 1 (\Sexpr{coefs_data %>% filter(case == "bad", feature == "Feature 1") %>% pull(b1) %>% round(3) %>% abs()}). Thus, this explainer model actually provides a misleading explanation that Feature 2 plays a more important role in the prediction of interest.}

%\kge{To demonstrate the effect of tuning parameters on LIME,we consider two different explainer models applied to the same prediction. Suppose a hypothetical black-box model is trained using one feature. Both plots in Figure1show the black-box model predictions versus the feature used to train the model.The diamond shaped point in the plots indicates the location of the prediction of interest to explain. A weight ridge regression with the black-box predictions as the response variable and standardized versions the features as predictor variables will be used as the explainer model in both scenarios, but the weights will vary. Note that the standardized features allow for direct comparisons of the model coefficients.}

\autoref{fig:figure-01} demonstrates a simple example in which the explanation quality varies drastically when different weights are used. Other concerns with LIME have been raised in the literature: \citet{laugel:2018} and \citet{molnar:2019} discuss the difficulty specifying a local region with LIME due to both  an unclear definition of what a "local region" is and, secondly, how to achieve an appropriate local region in LIME. \citet{laugel:2018} also present examples of LIME explanations that are clearly global and not local. A different concern about LIME is raised by \citet{alvarezmelis:2018}  pertaining to the robustness of explanations from LIME and other explainer models: they find that even small changes in predictor variables can lead to very different LIME explanations. Even the original authors of LIME,  \citet{ribeiro:2016}, acknowledge that if a linear model is used as the explainer, LIME relies on a linear approximation of the explainer model to the complex model and state "if the underlying model is highly non-linear even in the locality of the prediction, there may not be a faithful explanation".

 Without an assessment of the LIME explainer model, the user has to put trust in yet another black-box procedure to explain the complex black-box model. In this paper, we stress the importance of assessing the LIME explainer model. \hh{XXX This is where the paragraph has to go XXX} To do this, we lay out the set of claims about LIME made by \citet{ribeiro:2016}, and we provide a toolkit of visual diagnostics  for the assessment of these claims: (1) \emph{explanation scatterplots}, (2) \emph{feature heatmaps}, (3) \emph{assessment metric plots}. Using these visual diagnostics, we demonstrate problematic examples of LIME explanations where the claims about LIME are not met and the explanations are highly dependent on tuning parameter values without a clear way to identify a set of tuning parameter values that produce the "best" explanations.

@@

\citet{ribeiro:2016} provide an implementation of LIME in a Python package \cite{ribeiro:2020}. An adaption of the Python package in R has been implemented and made available by Thomas Lin-Pedersen \citep{pedersen:2020}. In both implementations, a ridge regression model is used as the explainer model, and  the user is asked to select various tuning parameter options: the number of features to return in the explanation, the simulation method, the feature selection method, and how the weights are computed. The discussion of parameter choices and implementation details of LIME in this paper are based on the R package. See Appendix~\ref{lime-details} for
an overview of the options available for the tuning parameters.

@@

@@

@
\caption{Plots of $x_2$ versus $x_1$ from the \data \  training (left) and testing (right) sets introduced in Section \ref{diagnostics}  with the true classification boundary  shown as a solid black line.  In the testing set, the open black circles identify cases misclassified by the random forest, and the diamond indicates a prediction of interest. }
\label{fig:figure-02}
\end{figure*}

@@

@
\caption{(left) Visualization from the \emph{lime} R package of a LIME explanation for the \data \ prediction of interest identified in \autoref{fig:figure-02}. (right) \emph{Explanation Scatterplot}.  An explanation scatterplot associated with the explanation  on the left. The four quantile bins over-simplify 
the relationship between the random forest predictions and the $x_1$ and $x_2$ values near the prediction of interest.}
\label{fig:figure-03}
\end{figure*}

@@

To demonstrate the visual diagnostics, we generate an example dataset that will be referred to as the \data. The \data \ contains \Sexpr{dim(sine_data)[1]} observations with  features of $x_1$ and $x_2$ independently sampled from Unif(\Sexpr{min}, \Sexpr{max}) distributions and $x_3$  sampled from a $\mbox{N}(0,1)$ distribution. A binary response variable $y$ is created using a  sine curve  such that
\begin{eqnarray}\label{eq:data}
  y=\begin{cases}
  \mbox{blue} & \mbox{ if } x'_2 > \sin\left(x'_1\right) \\
  \mbox{red} & \mbox{ if } x'_2 \le \sin\left(x'_1\right) \ .
  \end{cases}
\end{eqnarray}
where $x'_1=x_1\cos(\theta)-x_2\sin(\theta)$,  $x'_2=x_1\sin(\theta)+x_2\cos(\theta)$, and  $\theta=-0.9$. \sout{Then $y$ is defined as
\begin{eqnarray}\label{eq:data}
  y=\begin{cases}
  \mbox{blue} & \mbox{ if } x'_2 > \sin\left(x'_1\right) \\
  \mbox{red} & \mbox{ if } x'_2 \le \sin\left(x'_1\right) \ .
  \end{cases}
\end{eqnarray}}
Note that due to the creation of $y$ in this manner, $y$ is dependent on $x_1$ and $x_2$ and independent of $x_3$. The dataset is randomly divided into training and testing sets of \Sexpr{dim(sine_data_train)[1]} and \Sexpr{dim(sine_data_test)[1]} observations, respectively. A random forest is fit to the training set using the R package \emph{randomForest} (version \Sexpr{packageVersion("randomForest")}) \citep{liaw:2002} with the default settings and is applied to the test set to obtain predictions.

@@

Since 4-quantile-bins are used to simulate data, LIME converts continuous predictor variables to indicator variables identifying whether the variable value falls in the same quantile bin as the prediction of interest or not. The indicator variables are used as the features in the ridge regression explainer model. The features included in the visualization are the ones selected by feature selection, the lengths of the bars represent the coefficients from the ridge regression associated with the indicator variables, while the color of a bar denotes the sign of the coefficient. This explainer model suggests that a random forest prediction of `blue' for this observation is mostly supported by the prediction of interest having a value of $x_2$ that is greater than \Sexpr{str_split(sine_poi_lime_default$feature_desc[[1]], pattern = " ")[[1]][1]} but somewhat contradicted by the prediction of interest having a value of $x_1$ that is greater than \Sexpr{str_split(sine_poi_lime_default$feature_desc[[2]], pattern = " ")[[1]][1]} and less than or equal to \Sexpr{str_split(sine_poi_lime_default$feature_desc[[2]], pattern = " ")[[1]][5]}. \kgc{I decided to leave the values here. I actually prefer it this way myself, but I cut out a lot of text from this subsection in other ways.}\hh{XXX }

@@

For a further assessment of the explainer model, we turn to an \emph{explanation scatterplot}: a visual diagnostic for assessing the LIME claims of locality and fidelity for an individual explanation by juxtaposing the complex and explainer models in one plot. The format of an explanation scatterplot depends on the LIME simulation method. We introduce the explanation scatterplot here under the \emph{lime} R package default method of 4-quantile-bins. Appendix \ref{exp-scatter_plus} addresses the explanation scatterplot formats in other LIME simulation scenarios.

The explanation scatterplot applies the concept of plotting the model in the data space discussed in \citet{wickham:2015}. The plot is built by plotting the LIME simulated data for the top two features identified by the explanation in a scatterplot and coloring these points by the predictions from the complex model. The point size represents the weight assigned by LIME. In order to show the LIME results for the observation of interest, lines are drawn on top of the points  representing the boundaries of the indicator variables used to fit the explainer model. The line color  denotes whether LIME indicates that a feature supports or contradicts a class prediction.

@@

First, consider the claim of localness. The weights decay relatively slowly outside of the intersection of the 2-quantile-bin suggesting that the LIME explanation is highly influenced by points outside of the bins containing the prediction of interest. However, it is difficult to say if the claim of localness has been violated, because a definition of a local region is not specified. Depending on what region a viewer considers to be local, an argument could be made in favor of or against a violation of localness. The explanation scatterplot raises awareness of the unclear definition of a local region with LIME. Nevertheless, it is possible to say that the weights assigned to the simulated data do not capture  the local region identified by the intersection of the quantile bins.

Now, consider the claim of faithfulness: It can be said that the majority of the points in the $x_2$ quantile support a prediction of blue, which is captured by the bar supporting a prediction of blue in the LIME explanation, and a similar statement can be made about the $x_1$ quantile-bin. These statements validate the explanation produced by LIME. However, the explanation scatterplot plot shows that the random forest performs well at capturing the sine curve classification boundary by creating various sized rectangles consisting of predictions with similar probabilities, which the LIME explanation does not pick up on since the bins are created without information about the random forest predictions. Thus, LIME does provide an explanation for the prediction, but it is a very poor explanation in terms of faithfulness to the random forest prediction regions.

@@

This example explanation only includes two features. In situations where more than two features are included in a LIME explanation, the explanation scatterplots can be extended to a generalized pairs plot \citep{emerson:2013} that includes all pairwise combinations of features. Generalized pairs plots (and scatterplot matrices in general) have diminishing value when the number of features increase \citep{jensen:2011} \citep{sweller:2011}. Machine learning models are commonly fit using a large number of features, and therefore, a generalized pairs plot of explanation scatterplots for all features would be ineffective. However, when applying LIME, the user selects the number of features to return in the explanation. In the \emph{lime} R package, \citet{pedersen:2020} encourage users to select less than 10 features. As long as a small number of features are returned in the LIME explanation,  it is feasible to use a generalized pairs plot of explanation scatterplots. An example is shown in Section~\ref{bullet-assess-ex}.

@@

@
\caption{Example bullet signatures. The  signatures are from  the same land and therefore have very similar patterns. }
\label{fig:figure-07}
\end{figure}

@@

\begin{figure*}[!thp]
<<figure-11, out.width = '6.25in', fig.width = 13, fig.height = 8.5, message = FALSE>>=

@@

\section{Additional Bullet Matching Explanation Scatterplots} \label{bullets-plus}
